\mytitle

### 2016-10-02 pdocs/thesis/article-section-2.tex
Fixes in article based on Timo's comments

% pdf page 33, notation in typical ODEs
\begin {table}[htb!]
\begin{center}
\begin{tabular}{|l| l|}
\hline
{\bf Symbol} & {\bf Description} \\  \hline
$\vec{r}_i$ & position of center of mass for body i  \\ \hline
$\vec{q}_i$ & orientation for body i as quaternion $\lbrack s_i, x_i, y_i, z_i \rbrack ^T $ \\
\hline
$\vec{p}_i$ & contact or joint point i  \\ \hline
$\vec{r}_{ki}$ & $\vec{p}_k - \vec{r}_i$  \\ \hline
$\vec{s}$ & $\lbrack \vec{r}_1, \vec{q}_1,...,\vec{r}_n, \vec{q}_n \rbrack ^T $\\ \hline
$Q_i$ & \begin{tabular}{@{}c}
rotation of quaternion $\vec{q}_i$
as matrix \\ where 
$\frac{1}{2}\vec{\omega}_i \vec{q}_i=Q_i \vec{\omega}_i$
\end{tabular}
   $
\frac{1}{2} \left[ \begin{array}{ccc}
-x_i & -y_i & -z_i \\
s_i & z_i & -y_i \\
-z_i & s_i & x_i \\
y_i & -x_i & s_i
\end{array} \right]
$
\\ \hline
$S$ & 
\begin{tabular}{@{}c}
generalized transformation matrix \\
$ S \in \mathbb{R}^{7n \times 6n}$
\end{tabular}
 $ \left[ \begin{array}{ccccc}
1 &  &  & & 0 \\
 & Q_i  \\
 & & \ddots  \\
 & & & 1 \\
0 & & & & Q_n 
\end{array} \right]
$
\\ \hline
$\vec{v}_i$ & linear velocity of  center of mass for body i   \\ \hline
$\vec{\omega}_i$ & angular velocity of center of mass for body i  \\ \hline
$\vec{u}$ & $\lbrack \vec{v}_1, \vec{\omega}_1,...,\vec{v}_n, \vec{\omega}_n \rbrack ^T $\\ \hline
$M$ &
\begin{tabular}{@{}c}
 generalized mass matrix \\
$ M \in \mathbb{R}^{6n \times 6n}$  
\end{tabular}
$
\left[ \begin{array}{ccccc}
m_i 1 &  &  & & 0 \\
 & I_1  \\
 & & \ddots  \\
 & & & m_n 1 \\
0 & & & & I_n 
\end{array} \right]
$
\\ \hline
$I_i$ & inertia tensor for body i  \\ \hline
$C$ & contact condition matrix  $ C \in \mathbb{R}^{6n \times 3K}$ \\ \hline
$N$ & contact normal matrix  $ N \in \mathbb{R}^{3K \times K}$ \\ \hline
\end {tabular}
\end{center}
\caption {Nomenclature for equations of motion} \label{tab:eom}
\end {table}

@@

% pdf page 33, notation in typical ODEs
\begin {table}[htb!]
\begin{center}
\begin{tabular}{|l| l|}
\hline
{\bf Symbol} & {\bf Description} \\  \hline
$J_{contact}$ & Jacobian matrix for contacts  \\ \hline
$\lambda_{contact}$ & vector of lagrange multipliers for contacts  \\ \hline
$J_{joint}$ & Jacobian matrix for joints  \\ \hline
$\lambda_{joint}$ & vector of lagrange multipliers for joints  \\ \hline
\end {tabular}
\end{center}
\caption {Additional terms  for generalized equations of motion} \label{tab:eom-g}
\end {table}

@@

\begin {table}[htb!]
\begin{center}
\begin{tabular}{|c| l| l|}
\hline
{\bf Parameter} & {\bf Description} & {\bf btConstraintInfo2 pointer}\\  \hline
$J_1, \Omega_1$ & jacobian & m\_J1linearAxis, m\_J1angularAxis \\
$J_2, \Omega_2$ & & m\_J2linearAxis, m\_J2angularAxis \\ \hline
$\vec{v}$ & linear velocity & \\ \hline
$\vec{\omega}$ & angular velocity & \\ \hline
$\vec{c}$        &  right side vector   & m\_constraintError \\ \hline
$C$  & constraint force mixing & cfm \\  \hline
$\vec{\lambda}$ & constraint force &  \\ \hline
$\vec{l}$ & low limit for constraint force & m\_lowerLimit \\ \hline
$\vec{h}$ & high limit for constraint force & m\_upperLimit \\ \hline
\end {tabular}
\end{center}
\caption {Constraint parameters} \label{tab:constraintParameters} 
\end {table}

@@

\begin{figure}[htb!]
\centering
\begin{tikzpicture}
\coordinate (S) at (2.5,5);
\draw (0,5) -- (4,5) ;
\draw (0,0) -- (4,0) ;
\draw (2,0) -- (2,5) ;
\draw (1.5,0) -- (S); 
\node[above] at (S) [align=center] {\large{$\sigma<f_y$}};
\node[anchor=west] at (-2,3) {
\begin{tabular}{l}
Under elastic load\\
stress increases\\
linearly from zero\\
at neutral axis to\\
maximum value at \\
surface of body
\end{tabular}
};
\end{tikzpicture}
\hspace{1cm}
\begin{tikzpicture}
\coordinate (S) at (3,5);
\draw (0,5) -- (4,5) ;
\draw (0,0) -- (4,0) ;
\draw (2,0) -- (2,5) ;
\draw (1,0) -- (1,2.5) -- (3,2.5) -- (S); 
\node[above] at (S) [align=center] {\large{$\sigma=f_y$}};
\node[anchor=west] at (3,3) {
\begin{tabular}{l}
Under plastic load\\
stress is at yield\\
level over full\\
cross section
\end{tabular}
};
\end{tikzpicture}
\caption{Axial stress distribution over cross section for bending under elastic and plastic loads.}
\label{fig:wp}
\end{figure}

@@

Maximum moments acting around axis perpendicular to $\vec{r}_{anc}^{\,i} $
are integrals of perpdendicular distance 
and yield stress $f_y$ as given for the moment around x-axis and,
correspondingly, moment around z-axis.

>>>>

% pdf page 33, notation in typical ODEs
\begin {table}[htb!]
\caption {Nomenclature for equations of motion}
\label{tab:eom}
\begin{center}
\begin{tabular}{|l| l|}
\hline
{\bf Symbol} & {\bf Description} \\  \hline
$\vec{r}_i$ & position of center of mass for body i  \\ \hline
$\vec{q}_i$ & orientation for body i as quaternion $\lbrack s_i, x_i, y_i, z_i \rbrack ^T $ \\
\hline
$\vec{p}_i$ & contact or joint point i  \\ \hline
$\vec{r}_{ki}$ & $\vec{p}_k - \vec{r}_i$  \\ \hline
$\vec{s}$ & $\lbrack \vec{r}_1, \vec{q}_1,...,\vec{r}_n, \vec{q}_n \rbrack ^T $\\ \hline
$Q_i$ & \begin{tabular}{@{}c}
rotation of quaternion $\vec{q}_i$
as matrix \\ where 
$\frac{1}{2}\vec{\omega}_i \vec{q}_i=Q_i \vec{\omega}_i$
\end{tabular}
   $
\frac{1}{2} \left[ \begin{array}{ccc}
-x_i & -y_i & -z_i \\
s_i & z_i & -y_i \\
-z_i & s_i & x_i \\
y_i & -x_i & s_i
\end{array} \right]
$
\\ \hline
$S$ & 
\begin{tabular}{@{}c}
generalized transformation matrix \\
$ S \in \mathbb{R}^{7n \times 6n}$
\end{tabular}
 $ \left[ \begin{array}{ccccc}
1 &  &  & & 0 \\
 & Q_i  \\
 & & \ddots  \\
 & & & 1 \\
0 & & & & Q_n 
\end{array} \right]
$
\\ \hline
$\vec{v}_i$ & linear velocity of  center of mass for body i   \\ \hline
$\vec{\omega}_i$ & angular velocity of center of mass for body i  \\ \hline
$\vec{u}$ & $\lbrack \vec{v}_1, \vec{\omega}_1,...,\vec{v}_n, \vec{\omega}_n \rbrack ^T $\\ \hline
$M$ &
\begin{tabular}{@{}c}
 generalized mass matrix \\
$ M \in \mathbb{R}^{6n \times 6n}$  
\end{tabular}
$
\left[ \begin{array}{ccccc}
m_i 1 &  &  & & 0 \\
 & I_1  \\
 & & \ddots  \\
 & & & m_n 1 \\
0 & & & & I_n 
\end{array} \right]
$
\\ \hline
$I_i$ & inertia tensor for body i  \\ \hline
$C$ & contact condition matrix  $ C \in \mathbb{R}^{6n \times 3K}$ \\ \hline
$N$ & contact normal matrix  $ N \in \mathbb{R}^{3K \times K}$ \\ \hline
\end {tabular}
\end{center}
\end {table}

@@

% pdf page 33, notation in typical ODEs
\begin {table}[htb!]
\caption {Additional terms  for generalized equations of motion}
\label{tab:eom-g}
\begin{center}
\begin{tabular}{|l| l|}
\hline
{\bf Symbol} & {\bf Description} \\  \hline
$J_{contact}$ & Jacobian matrix for contacts  \\ \hline
$\lambda_{contact}$ & vector of lagrange multipliers for contacts  \\ \hline
$J_{joint}$ & Jacobian matrix for joints  \\ \hline
$\lambda_{joint}$ & vector of lagrange multipliers for joints  \\ \hline
\end {tabular}
\end{center}
\end {table}

@@

\begin {table}[htb!]
\caption {Constraint parameters}
\label{tab:constraintParameters} 
\begin{center}
\begin{tabular}{|c| l| l|}
\hline
{\bf Parameter} & {\bf Description} & {\bf btConstraintInfo2 pointer}\\  \hline
$J_1, \Omega_1$ & jacobian & m\_J1linearAxis, m\_J1angularAxis \\
$J_2, \Omega_2$ & & m\_J2linearAxis, m\_J2angularAxis \\ \hline
$\vec{v}$ & linear velocity & \\ \hline
$\vec{\omega}$ & angular velocity & \\ \hline
$\vec{c}$        &  right side vector   & m\_constraintError \\ \hline
$C$  & constraint force mixing & cfm \\  \hline
$\vec{\lambda}$ & constraint force &  \\ \hline
$\vec{l}$ & low limit for constraint force & m\_lowerLimit \\ \hline
$\vec{h}$ & high limit for constraint force & m\_upperLimit \\ \hline
\end {tabular}
\end{center}
\end {table}

@@

\begin{figure}[htb!]
\centering
\begin{tikzpicture}
\coordinate (S) at (2.5,5);
\draw (0,5) -- (4,5) ;
\draw (0,0) -- (4,0) ;
\draw (2,0) -- (2,5) ;
\draw (1.5,0) -- (S); 
\node[above] at (S) [align=center] {\large{$\sigma<f_y$}};
\node[anchor=west] at (-2,3) {
\begin{tabular}{l}
Under elastic load\\
stress increases\\
linearly from zero\\
at neutral axis to\\
maximum value at \\
surface of body
\end{tabular}
};
\end{tikzpicture}
\hspace{1cm}
\begin{tikzpicture}
\coordinate (S) at (3,5);
\draw (0,5) -- (4,5) ;
\draw (0,0) -- (4,0) ;
\draw (2,0) -- (2,5) ;
\draw (1,0) -- (1,2.5) -- (3,2.5) -- (S); 
\node[above] at (S) [align=center] {\large{$\sigma=f_y$}};
\node[anchor=west] at (3,3) {
\begin{tabular}{l}
Under fully plastic load\\
stress is at yield\\
level over full\\
cross section
\end{tabular}
};
\end{tikzpicture}
\caption{Axial stress distribution over cross section for bending under elastic and fully plastic loads.}
\label{fig:wp}
\end{figure}

@@

Maximum moments acting around axis perpendicular to $\vec{r}_{anc}^{\,i} $
are integrals of perpdendicular distance 
and yield stress $f_y$ as given for the moment around $x$-axis 
and moment around $z$-axis, respectively.

### 2016-10-02 pdocs/thesis/article-section-3.tex
article tuning
Fixes in article based on Timo's comments

System has only one dynamic rigid body which is three meters high block of 0.04 \% steel enforced concrete. 
Cross section is one square meter.
Constraint is set so that connecting frame is at top of block. 
In this case frame could be anywhere but if multiple bodies are involved, connecting frame must be defined
so that it reflects scenario under investigation.
Concrete density is $2000 \frac{kg}{m^3}$ and steel density is $7800 \frac{kg}{m^3}$. 
Steel is assumed to handle load in elastic-plastic case. Yield stress of steel is 200 MPa.
Gravity is $10 \frac{m}{s^2}$. Simulation step is $\frac{1}{60} s$ and 10 
iterations are done for single step.
Body is 1.5 meters above rigid ground. In unconstrained case body will hit hit ground at t=0.548 s 
($\sqrt{2 \cdot 1.5 / 10}$).

Single simulation step is done using following substeps. 
In this work, changes are made only to constraint setup and update actions.
\begin{enumerate}
\item Apply gravity to each non static body. 
This step makes programming easier as otherwise each program should add forces due to gravity in each step. 
In this case, $\vec{f}_{ext}$ in  Equation \ref{eq:eomV2} gets added by $\lbrace{0,-60000,0}\rbrace$.
\item Predict unconstrained motion for each body controlled by physics simulation.
Static and kinematic bodies are not processed in this step.
Prediction is done based on current linear and angular velocities of the body.
\item Predict contacts. In this phase, continuous collision detection is done based on simplified objects. 
Each rigid body is represented by sphere geometry and contact prediction is done if body moves more than given threshold value during simulation step. Continuous collision detection is configured for each body. 
Typical scenario that requires continuous collision detection is fast moving bodies that would otherwise go through walls.
\item Perform discrete collision detection. All overlapping bodies are processed and manifoldPoints are created for
each detected contact at end of simulation step. 
In unconstrained case, contact is detected after $t$=0.55 s if continuous collision detection is not used and manifolds distance gets negative value (-0.058 m).
\item Calculate simulation islands(groups). Bodies that are near each other based on contact prediction or discrete collision detection
 or connected with constraints are grouped in same group.
\item Solve constraints. Both contact and other constraints are processed in this step. 
 This step is subdivided to setup, iterations and finish phases. 
In setup phase constraints are queried for positional errors for calculation of $c$ in
Equation \ref{eq:constraintEquation} and maximum and minimum impulses in Equations \ref{eq:lambdaLow} and 
\ref{eq:lambdaHigh}.
In finish phase constraint forces are calculated if requested and velocities of bodies are updated.
\item Integrate transforms using velocities modified in previous step.
\item Update actions (callbacks) are called. 
In elastic-plastic case, plasticity is summed and equilibrium point of elastic part is updated if maximum force or moment is exceeded.
\item Activation state of bodies is updated. 
To avoid extra calculation bodies are as default put to sleeping state if linear and angular velocities of body are less than
given threshold values (default 0.8 and 1.0) longer than set time limit (2.0).
\end{enumerate}

@@

\begin {table}[htb!]
\begin{center}
\begin{tabular}{|l| l|l| l|l|l|l|l|}
\hline
{\bf Time} & 
{\bf Location} &
{\it velError} & {\it penetration} & {\it posError} & {\it rhs} &
{\bf Velocity} & 
{\bf Impulse} \\  \hline
0.017 &  0 & & & &  &-0.17 & 0 \\  \hline
0.550 &  -1.558 & & & & & -5.5 & 0 \\  \hline
0.567 &  -1.511 & 5.67 &-0.058 &2.8 &  21270 & 0.01 & 34000 \\  \hline
0.583 &  -1.502 & 0.14 &-0.011 & 0.54& 2570  & 0.55 & 420 \\  \hline
0.600 &  -1.496 & -0.38&-0.002 & 0.1  & -1000& 0.38 & 0 \\  \hline
0.617 &  -1.492 &-0.44 & 0.004 & 0     & -1600& 0.22 & 0 \\  \hline
0.717 &  -1.497 &0.004-0.08  &-0.0003-0.001 &0-0.01 & 10-400 & -0.01 & 400 \\  \hline
0.817 &  -1.499 & & & & & -0.08 & 700 \\  \hline
0.917 &  -1.500 & & & & & -0.001 & 1000 \\  \hline
\end {tabular}
\end{center}
\caption {Simulation values for unconstrained case. 
For internal contact values typical values are shown
as number of contacts and values at contact points are different.} 
\label{tab:freeBlockValues} 
\end {table}

@@

\begin {table}[htb!]
\begin{center}
\begin{tabular}{|l| l|l| l|l|l|l|l|}
\hline
{\bf Time} & 
{\bf Location} &
{\it velError} & {\it penetration} & {\it posError} & {\it rhs} &
{\bf Velocity} & 
{\bf Impulse} \\  \hline
0.017 &  0 & & & &  &-0.17 & 0 \\  \hline
0.550 &  -1.500 & 3.5 & 0.033 & 0 & 21000& -2 & 21000 \\  \hline
0.567 &  -1.500 & 2.17 & 0 & 0  &  8100 & 0.01 & 13000 \\  \hline
0.583 &  -1.500 & 0.15 & 0 & 0 & 600  & 0 & 937 \\  \hline
0.600 &  -1.500 & 0.17 & 0 & 0 & 600  & 0 & 1040 \\  \hline
\end {tabular}
\end{center}
\caption {Simulation values for unconstrained case with continuous collision detection. } 
\label{tab:freeBlockValuesWithCcd} 
\end {table}

@@

\begin {table}[htb!]
\begin{center}
\begin{tabular}{|l|l| l| l|l|l|l|}
\hline
{\bf Time} & 
{\bf Location} &
{\it velError} & {\it posError} & {\it rhs} &
{\bf Velocity} & 
{\bf Impulse} \\  \hline
0.017 & -0.0028 &  & & & -0.17 & 0 \\  \hline
0.033 & -0.0022 & 0.33 & -0.033 & -2200 & 0.033 & 2200 \\  \hline
0.050 & -0.0018 & -0.13 & -0.027 & -960 & 0.027 & 960 \\  \hline
0.067 & -0.0014 &-0.14 & -0.021 & -970 & 0.021 & 970 \\  \hline
0.35... & 0 &-0.17 & $\approx$ 0 & -1000 &0.0 & 1000 \\  \hline
\end {tabular}
\end{center}
\caption {Constraint parameter values for fixed constraint} \label{tab:fixedBlockValues} 
\end {table}

@@

\begin {table}[htb!]
\begin{center}
\begin{tabular}{|l|p{10cm}|}
\hline
{\bf Parameter} & 
{\bf Description} 
\\ \hline
lowerLimit &  
minimum allowed translation or rotation
 \\  \hline
upperLimit &  
maximum allowed translation or rotation
 \\  \hline
springStiffness & elastic spring stiffness
 \\  \hline
enableSpring & defines if spring is active
 \\  \hline
springStiffnessLimited & should elastic behaviour be tuned to avoid instability
 \\  \hline
equilibriumPoint & should elastic behaviour be tuned to avoid instability
 \\  \hline
currentLimit & describes state of constraint \newline
 0: not limited \newline
 3: loLimit=hiLimit \newline 
 4: current value is between loLimit and HiLimit
 \\ \hline

\end {tabular}
\end{center}
\caption {Selected constraint parameters for elastic-plastic constraint 2} \label{tab:ep2Parameters} 
\end {table}

@@

\begin {table}[htb!]
\begin{center}
\begin{tabular}{|l|l| l| l|l|l|l|l|}
\hline
{\bf Time} & 
{\bf Location} &
{\it velError} & {\it posError} & {\it rhs} &
{\it velocity} & 
{\bf Impulse} & 
{\bf Plastic strain} \\  \hline
0.017 & -0.0028 &-0.17 & 0 & -1000 & -0.17 & 0 & 0 \\  \hline
0.033 & -0.0046 &-0.33 & 0 & -2000 & -0.11 &  1330 & 0.001 \\  \hline
0.050 & -0.0056 &-0.28 & 0 & -1670 & -0.056 &  1330 & 0.003 \\  \hline
0.067 & -0.0056 &-0.22 & 0 & -1340 &  -0.001&  1330 & 0.004\\  \hline
0.083... & -0.0056  & -0.17 & 0 & -1000 &  0.0&  1000 & 0.004\\  \hline
\end {tabular}
\end{center}
\caption {Constraint parameter values for elastic-plastic constraint} \label{tab:epBlockValues} 
\end {table}

@@

\begin {table}[htb!]
\begin{center}
\begin{tabular}{|l|l| l| l|l|l|l|l|}
\hline
{\bf Time} & 
{\bf Location} &
{\it velError} & {\it posError} & {\it rhs} &
{\bf Velocity} & 
{\bf Impulse} & 
{\bf Plastic strain} \\  \hline
0.017... &  0 & -0.17  & 0 & -1000 & 0       & 1000 & 0 \\  \hline
\end {tabular}
\end{center}
\caption {Constraint parameter values for elastic-plastic constraint 2} \label{tab:ep2BlockValues} 
\end {table}

\subsection{Charpy impact test}
Simulation of Charpy impact test was used to benchmark approach in more complicated scenario. 
Material is steel. Density is 7800 $\frac{kg}{m^{3}}$. Young’s modulus is 200 GPa.
Specimen dimensions are 10x10x55 mm with 2 mm notch in middle which is taken into account in calculations.
Support anvils initially have 40 mm open space between them. Their width is 40 mm. 
Hammer is 0.5 m wide and 0.25 m high, thickness is 0.02 m. Mass is 19.5 kg.
Hammer has 40 mm draft.
If specimen bends about 1.9 radians (108 degrees) it will go between anvils.
Expected energy loss is product of plastic moment of section, hinge angle needed for specimen to go through supports and 
yield stress of specimen. For hinge angle of 1.9 radians and yield stress of 400 MPa expected energy 
loss is 122 J.

@@

\begin {table}[htb!]
\begin{center}
\begin{tabular}{| c| c|l|}
\hline
{\bf Timestep[ms]} & {\bf Energy loss [J]} & {\bf Notes} \\ \hline
 0.2 &  170 & specimen penetrates hammer  \\ \hline
 0.1 &  120 & \\ \hline
 0.05 &  150 & \\ \hline
\end {tabular}
\end{center}
\caption {Energy loss for few timesteps} \label{tab:ep2ts} 
\end {table}

@@

\subsection{Demolisher}
Demolisher is  collection of possible scenarios that can be added to games to provide ductile joints between
rigid bodies.  Simulation can be done in real time on modest hardware. 
Figure \ref{fig:demolisher} shows three screenshots from simulation.

>>>>

System has only one dynamic rigid body which is three meters high block of 0.04 \% steel enforced concrete. 
Cross section is one square meter.
Constraint is set so that connecting frame is at top of block. 
In this case frame could be anywhere but if multiple bodies are involved, connecting frame must be defined
so that it reflects scenario under investigation.
Concrete density is $2000\ \frac{kg}{m^3}$ and steel density is $7800\ \frac{kg}{m^3}$. 
Steel is assumed to handle load in elastic-plastic case. Yield stress of steel is 200 MPa.
Gravity is $10\ \frac{m}{s^2}$. Simulation step is $\frac{1}{60}\ s$ and 10 
iterations are done for single step.
Body is 1.5 meters above rigid ground. In unconstrained case body will hit hit ground at t = 0.548 s 
($\sqrt{2 \cdot 1.5 / 10}$).

Single simulation step is performed using following substeps. 
In this work, changes are made only to constraint setup and update actions.
\begin{enumerate}
\item Apply gravity to each non static body. 
This step makes programming easier as otherwise each program should add forces due to gravity in each step. 
In this case, $\vec{f}_{ext}$ in  Equation \ref{eq:eomV2} gets added by $\lbrace{0,-60000,0}\rbrace$.
\item Predict unconstrained motion for each body controlled by physics simulation.
Static and kinematic bodies are not processed in this step.
Prediction is done based on current linear and angular velocities of the body.
\item Predict contacts. In this phase, continuous collision detection is done based on simplified objects. 
Each rigid body is represented by sphere geometry and contact prediction is done if body moves more than given threshold value during simulation step. Continuous collision detection is configured for each body. 
Typical scenario that requires continuous collision detection is fast moving bodies that would otherwise go through walls.
\item Perform discrete collision detection. All overlapping bodies are processed and manifoldPoints are created for
each detected contact at end of simulation step. 
In unconstrained case, contact is detected after $t$=0.55 s if continuous collision detection is not used and manifolds distance gets negative value (-0.058 m).
\item Calculate simulation islands(groups). Bodies that are near each other based on contact prediction or discrete collision detection
 or connected with constraints are grouped in same group.
\item Solve constraints. Both contact and other constraints are processed in this step. 
 This step is subdivided to setup, iterations and finish phases. 
In setup phase constraints are queried for positional errors for calculation of $c$ in
Equation \ref{eq:constraintEquation} and maximum and minimum impulses in Equations \ref{eq:lambdaLow} and 
\ref{eq:lambdaHigh}.
In finish phase constraint forces are calculated if requested and velocities of bodies are updated.
\item Integrate transforms using velocities modified in previous step.
\item Update actions (callbacks) are called. 
In elastic-plastic case, plasticity is summed and equilibrium point of elastic part is updated if maximum force or moment is exceeded.
\item Activation state of bodies is updated. 
To avoid extra calculation bodies are as default put to sleeping state if linear and angular velocities of body are less than
given threshold values (default 0.8 and 1.0) longer than set time limit (2.0).
\end{enumerate}

@@

\begin {table}[htb!]
\caption {Simulation values for unconstrained case. 
For internal contact values typical values are shown
as number of contacts and values at contact points are different.} 
\label{tab:freeBlockValues} 
\begin{center}
\begin{tabular}{|l| l|l| l|l|l|l|l|}
\hline
{\bf Time} & 
{\bf Location} &
{\it velError} & {\it penetration} & {\it posError} & {\it rhs} &
{\bf Velocity} & 
{\bf Impulse} \\  \hline
0.017 &  0 & & & &  &-0.17 & 0 \\  \hline
0.550 &  -1.558 & & & & & -5.5 & 0 \\  \hline
0.567 &  -1.511 & 5.67 &-0.058 &2.8 &  21270 & 0.01 & 34000 \\  \hline
0.583 &  -1.502 & 0.14 &-0.011 & 0.54& 2570  & 0.55 & 420 \\  \hline
0.600 &  -1.496 & -0.38&-0.002 & 0.1  & -1000& 0.38 & 0 \\  \hline
0.617 &  -1.492 &-0.44 & 0.004 & 0     & -1600& 0.22 & 0 \\  \hline
0.717 &  -1.497 &0.004-0.08  &-0.0003-0.001 &0-0.01 & 10-400 & -0.01 & 400 \\  \hline
0.817 &  -1.499 & & & & & -0.08 & 700 \\  \hline
0.917 &  -1.500 & & & & & -0.001 & 1000 \\  \hline
\end {tabular}
\end{center}
\end {table}

@@

\begin {table}[htb!]
\caption {Simulation values for unconstrained case with continuous collision detection. } 
\label{tab:freeBlockValuesWithCcd} 
\begin{center}
\begin{tabular}{|l| l|l| l|l|l|l|l|}
\hline
{\bf Time} & 
{\bf Location} &
{\it velError} & {\it penetration} & {\it posError} & {\it rhs} &
{\bf Velocity} & 
{\bf Impulse} \\  \hline
0.017 &  0 & & & &  &-0.17 & 0 \\  \hline
0.550 &  -1.500 & 3.5 & 0.033 & 0 & 21000& -2 & 21000 \\  \hline
0.567 &  -1.500 & 2.17 & 0 & 0  &  8100 & 0.01 & 13000 \\  \hline
0.583 &  -1.500 & 0.15 & 0 & 0 & 600  & 0 & 937 \\  \hline
0.600 &  -1.500 & 0.17 & 0 & 0 & 600  & 0 & 1040 \\  \hline
\end {tabular}
\end{center}
\end {table}

@@

\begin {table}[htb!]
\caption {Constraint parameter values for fixed constraint} 
\label{tab:fixedBlockValues} 
\begin{center}
\begin{tabular}{|l|l| l| l|l|l|l|}
\hline
{\bf Time} & 
{\bf Location} &
{\it velError} & {\it posError} & {\it rhs} &
{\bf Velocity} & 
{\bf Impulse} \\  \hline
0.017 & -0.0028 &  & & & -0.17 & 0 \\  \hline
0.033 & -0.0022 & 0.33 & -0.033 & -2200 & 0.033 & 2200 \\  \hline
0.050 & -0.0018 & -0.13 & -0.027 & -960 & 0.027 & 960 \\  \hline
0.067 & -0.0014 &-0.14 & -0.021 & -970 & 0.021 & 970 \\  \hline
0.35... & 0 &-0.17 & $\approx$ 0 & -1000 &0.0 & 1000 \\  \hline
\end {tabular}
\end{center}
\end {table}

@@

\begin {table}[htb!]
\caption {Selected constraint parameters for elastic-plastic constraint 2}
\label{tab:ep2Parameters} 
\begin{center}
\begin{tabular}{|l|p{10cm}|}
\hline
{\bf Parameter} & 
{\bf Description} 
\\ \hline
lowerLimit &  
minimum allowed translation or rotation
 \\  \hline
upperLimit &  
maximum allowed translation or rotation
 \\  \hline
springStiffness & elastic spring stiffness
 \\  \hline
enableSpring & defines if spring is active
 \\  \hline
springStiffnessLimited & should elastic behaviour be tuned to avoid instability
 \\  \hline
equilibriumPoint & should elastic behaviour be tuned to avoid instability
 \\  \hline
currentLimit & describes state of constraint \newline
 0: not limited \newline
 3: loLimit=hiLimit \newline 
 4: current value is between loLimit and HiLimit
 \\ \hline

\end {tabular}
\end{center}
\end {table}

@@

\begin {table}[htb!]
\caption {Constraint parameter values for elastic-plastic constraint} 
\label{tab:epBlockValues} 
\begin{center}
\begin{tabular}{|l|l| l| l|l|l|l|l|}
\hline
{\bf Time} & 
{\bf Location} &
{\it velError} & {\it posError} & {\it rhs} &
{\it velocity} & 
{\bf Impulse} & 
{\bf Plastic strain} \\  \hline
0.017 & -0.0028 &-0.17 & 0 & -1000 & -0.17 & 0 & 0 \\  \hline
0.033 & -0.0046 &-0.33 & 0 & -2000 & -0.11 &  1330 & 0.001 \\  \hline
0.050 & -0.0056 &-0.28 & 0 & -1670 & -0.056 &  1330 & 0.003 \\  \hline
0.067 & -0.0056 &-0.22 & 0 & -1340 &  -0.001&  1330 & 0.004\\  \hline
0.083... & -0.0056  & -0.17 & 0 & -1000 &  0.0&  1000 & 0.004\\  \hline
\end {tabular}
\end{center}
\end {table}

@@

\begin {table}[htb!]
\caption {Constraint parameter values for elastic-plastic constraint 2} 
\label{tab:ep2BlockValues} 
\begin{center}
\begin{tabular}{|l|l| l| l|l|l|l|l|}
\hline
{\bf Time} & 
{\bf Location} &
{\it velError} & {\it posError} & {\it rhs} &
{\bf Velocity} & 
{\bf Impulse} & 
{\bf Plastic strain} \\  \hline
0.017... &  0 & -0.17  & 0 & -1000 & 0       & 1000 & 0 \\  \hline
\end {tabular}
\end{center}
\end {table}

\subsection{Charpy impact test}
Simulation of Charpy impact test was used to benchmark approach in more complicated scenario. 
Material is steel and density is 7800 $\frac{kg}{m^{3}}$. Young’s modulus is 200 GPa.
Specimen dimensions are 10x10x55 mm with 2 mm notch in middle which is taken into account in calculations.
Support anvils initially have 40 mm open space between them. Their width is 40 mm. 
Hammer is 0.5 m wide and 0.25 m high, thickness is 0.02 m. Mass is 19.5 kg.
Hammer has 40 mm draft.
If specimen bends about 1.9 radians (108 degrees) it will go between anvils.
Expected energy loss is product of plastic moment of section, hinge angle needed for specimen to go through supports and 
yield stress of specimen. For hinge angle of 1.9 radians and yield stress of 400 MPa expected energy 
loss is 122 J.

@@

\begin {table}[htb!]
\caption {Energy loss for few timesteps}
\label{tab:ep2ts} 
\begin{center}
\begin{tabular}{| c| c|l|}
\hline
{\bf Timestep[ms]} & {\bf Energy loss [J]} & {\bf Notes} \\ \hline
 0.2 &  170 & specimen penetrates hammer  \\ \hline
 0.1 &  120 & \\ \hline
 0.05 &  150 & \\ \hline
\end {tabular}
\end{center}
\end {table}

@@

\subsection{Demolisher}
Demolisher is  collection of possible scenarios that can be added to games to provide ductile joints between
rigid bodies.  Rigid concrete bodies are connected by steel enforcement.
Concrete density is 2000 $\frac{kg}{m^3}$ and steel density is 7800 $\frac{kg}{m^3}$. 
Steel is assumed to handle load. Yield stress of steel is 200 MPa.
Gravity is 10  $\frac{m}{s^2}$. Simulation step is $\frac{1}{60} $ s and 10 
iterations are done for single step. Multiple simulation steps are done for each render if needed.

Simulation can be done in real time on modest hardware. 
Figure \ref{fig:demolisher} shows three screenshots from simulation.
